
# Predictive Model Training and Testing

## Objective
The objective of this project is to train a model and then test the model's ability to make predictions.  The model will be trained using both linear and nonlinear regression techniques, and the predictions made by the model will be compared to the actual values in the testing set to evaluate the model's performance.

## Tools Used
- R
- R Markdown

## Concepts Used


## Workflow
### Part 1: Load data, libraries, and define functions for later use
```{r}
rm(list = ls())
setwd('C:/Users/Max/Documents/MSBA/MSBA Fall B/MKT 436R/Assignment 3')
df <- read.csv('Student Data 5.csv')
```
Load packages
```{r}
library(earth)
```
Preserve raw data and rename table to an intuitive name
```{r}
approval_data <- data.frame(df)
```
Create training data set, made up of a random 80% of the total data set, and validation data set using the remaining data
```{r}
set.seed(3161)

is_training <- runif(nrow(approval_data)) < .8
training_data <- subset(approval_data, is_training)

validation_data <- subset(approval_data, !is_training)
```

Create RMSE Function to evaluate the accuracy of the model
```{r}
get_detail_RMSE <- function(model){
  actualY <- validation_data$card
  predictedY <- predict(model, validation_data)
  return(mean((actualY - predictedY)^2)^.5)
}
```
### Part 2: Creating the first predictive model
#### Part 2a: Preliminary Investigation
Run a MARS model so the plotmo function can be used to get more information to assist in the creation of models
```{r}
earth1 <- earth(card ~ reports + age + income + share + expenditure + owner + selfemp + dependents + months + majorcards + active, data = training_data)
plotmo(earth1)
```

The plotmo function indicates that share and reports appear to be parabolic.  These parameters will be squared in the first regression model to account for this relationship
```{r}
model1.1 <- lm(card ~ reports^2 + age + income + share^2 + expenditure + owner + selfemp + dependents + months + majorcards + active, data = training_data)
get_detail_RMSE(model1.1)
```
#### Part 2b: Refining the model using linear regression
Add an interaction between the non-linear terms seen in the plotmo of earth1 from step 2
```{r}
model1.2 <- lm(card ~ (reports^2 * share^2) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + share^2 + active, data = training_data)
model1.3 <- lm(card ~ (reports^2 * share^2) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), data = training_data)
model1.4 <- lm(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), data = training_data)

get_detail_RMSE(model1.2)
get_detail_RMSE(model1.3)
get_detail_RMSE(model1.4)
```
The accuracy of adding these terms improves the predictive capabilities of the model so they will be kept in future models
#### Part 2c: Refining the model using the MARS model
```{r}
# Now lets begin using the MARS modelon our best regression so far (model1B.4)
model1.5 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), data = training_data)
get_detail_RMSE(model1.5)
```
Implementing the MARS model increases the performance of the modelso this will be kept in future models.

#### Part 2d: Refining the model by applying the degree argument
```{r}
model1.6 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), degree = 2, data = training_data)
get_detail_RMSE(model1.6)
```
The degree term improves performance so this will be kept in future models.

#### Part 2e: Refining the model by tweaking the thresh term
```{r Part 2e}
model1.7 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
model1.8 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0.1, data = training_data)
model1.9 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + owner + selfemp + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0.01, data = training_data)

get_detail_RMSE(model1.7)
get_detail_RMSE(model1.8)
get_detail_RMSE(model1.9)
```
Thresh being equl to 0 and degree being equal to 2 yields the best results so model1.7 will be built upon moving forward
#### Part 2f: Visualize each parameter's relationship to card to further refine model
Visualize model1.7 to identify any other potential interactions
```{r}
plotmo(model1.7)
```
Additional interactions can be seen which will be added to the next model iteration.

```{r}
model1.10 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
get_detail_RMSE(model1.10)
plotmo(model1.10)
```
The visualization of model1.10 indicates that there are additional interactions that can be tested out in a new model
```{r}
model1.11 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + (reports^2 * dependents) + (owner * dependents) + months + (majorcards * owner) + (majorcards * selfemp) + (share^2 * active), degree = 2, thresh = 0, data = training_data)
get_detail_RMSE(model1.11)
```
Adding these interactions improved the model's performance so they will be kept.

Try adding factor terms to owner, selfemp, and majorcard to see if this improves performance further.
```{r}
model1.12 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * factor(owner)) + (factor(owner) * share^2) + (factor(selfemp) * factor(owner)) + dependents + months + factor(majorcards) + (share^2 * active), degree = 2, thresh = 0, data = training_data)

get_detail_RMSE(model1.12)
```
This didnt help so again model1.11 will continue to be used moving forward.

The last step is to try deleting terms one by one to see if that increases the model's performance
```{r}
model1.13 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
model1.14 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
model1.15 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
model1.16 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
model1.17 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)

get_detail_RMSE(model1.13)
get_detail_RMSE(model1.14)
get_detail_RMSE(model1.15)
get_detail_RMSE(model1.16)
get_detail_RMSE(model1.17)
```

Model 15 did better here so now expense will be dropped and poly will be added to age to see if any improvement can be yielded.

```{r}
model1.18 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + poly(age,3) + income + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
get_detail_RMSE(model1.18)
```
The improvement here was very small so model1.18 the final model




### Part 3: Validating the model
K-fold cross validation will be used on the best models to double check and make sure model1.18 is the best.  There are several models with near identical RMSE scores so those ones should be tested as well
```{r}
nFold = 5
model_performance <- matrix(NA, nFold, 3)

for(fold in 1:nFold){
  val_num <- floor(runif(nrow(approval_data))*nFold) + 1
  
  
  training_data <- subset(approval_data, val_num != fold)
  validation_data <- subset(approval_data, val_num == fold)
  
  model1.10 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
  model1.11 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + age + income + expenditure + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + (reports^2 * dependents) + (owner * dependents) + months + (majorcards * owner) + (majorcards * selfemp) + (share^2 * active), degree = 2, thresh = 0, data = training_data)
  model1.18 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + poly(age,3) + income + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = training_data)
  
  valid1 = mean((validation_data$card -
                   predict(model1.10,validation_data))^2)^.5
  valid2 = mean((validation_data$card -
                   predict(model1.11,validation_data))^2)^.5
  valid3 = mean((validation_data$card -
                   predict(model1.18,validation_data))^2)^.5
  model_performance[fold, ] <- c(valid1, valid2, valid3)
}

colMeans(model_performance)
```

After validation, it can be seen that model1.10 is actually the best so the final step is to train the model on all of the data available

```{r}
model1 <- earth(card ~ (reports^2 * share^2) + (reports^2 * active) + poly(age,3) + income + (reports^2 * owner) + (owner * share^2) + (selfemp * owner) + dependents + months + majorcards + (share^2 * active), degree = 2, thresh = 0, data = approval_data)
```


## Outcomes & Conclusions Made